### 2016-04-18 thesis/intro/intro.tex
tmp commit

\chapter{Introduction}

\section{Obesity}

General intro to obesity\\
Implications of obesity -- disease associated with it and/or cancer\\

\section{Cancer}

What is cancer\\

\subsection{Hallmarks of Cancer}

What is it caused by (11 hallmarks of cancer)\\

\subsection{Relationship between obesity}

Why is it important in obesity context\\

\section{Obesity Associated Genetic Signatures}

Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}\\

\section{Aim Of The Project}

>>>>

\chapter{Introduction}
\label{ch:intro}

\section{Obesity}
\label{sec:obesity}

General intro to obesity.
Implications of obesity -- disease associated with it and/or cancer.

\section{Cancer}
\label{sec:cancer}

What is cancer.

\subsection{Hallmarks of Cancer}
\label{subsec:cancerhallmarks}

What is it caused by (11 hallmarks of cancer).

\subsection{Relationship between obesity}
\label{subsec:obsbackground}

Why is it important in obesity context.

\section{Obesity associated genetic signatures}
\label{sec:obsgene}

Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}.

\section{Aim of the project}
\label{sec:aim}

This research project aims to determine whether gene expression signatures exist that are specific to obesity across multiple different cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in clinical settings in the future.

### 2016-04-18 thesis/methods/methods.tex
tmp commit

\chapter{Methods}

\section{Body Mass Index (BMI)}

\subsection{BMI Calculation}

BMI was calculated from the height and weight data of the samples using the following equation:\\
\begin{equation}
    BMI = \frac{weight (kg)}{Height^2(m^2)}
\end{equation}

\subsection{BMI Classification}

Samples were classified based on the WHO definition:
\begin{table}[h]
    \caption{WHO defined BMI classification}\\
    \label{tab:who-bmi-class}
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \bfseries {Classification} & \bfseries {BMI}\\
            \hline
            Underweight & \textless 20.0\\
            Normal weight/lean & 20.0$\sim$24.9\\
            Overweight & 25.0$\sim$29.9\\
            Obese & \textgreater 30.0\\
            \hline
        \end{tabular}
    \end{center}
\end{table}

\section{Download Data from publicly available source}

@@

\subsection{The Cancer Genome Atlas (TCGA)}

@@

\subsection{International Cancer Genome Consortium (ICGC)}

RNA-seq data for the eight cancer types with sample BMI information were downloaded from the ICGC database.

\section{Validation of methodology used by \citet{Creighton2012}}

\subsection{Raw data from \citet{Creighton2012} paper}

@@

\subsection{Normalisation of \citet{Creighton2012} data}

@@

\subsection{Differential gene expression analysis}

(mention WHO classification of obese/overweight/lean)\\
(also mention limma package)

\subsection{Singular Value Decomposition}

\section{Applying metagene to other cancer types}

\subsection{Transformation matrix}

\section{Pathway Enrichment Analysis}

>>>>

\chapter{Methods}
\label{ch:methods}

\section{Body Mass Index (BMI)}
\label{sec:bmi}

\subsection{BMI Calculation}
\label{subsec:bmicalc}

BMI was calculated from the height and weight data of the samples using the following equation:

\begin{equation}
    \label{eq:bmicalc}
    BMI = \frac{weight (kg)}{Height^2(m^2)}\\
\end{equation}

\subsection{BMI Classification}
\label{subsec:bmiclassification}

Samples were classified based on the WHO definition, as shown in Table~\ref{tab:whobmiclass}.
\begin{table}[hb]
    \caption{WHO defined BMI classification}
    \label{tab:whobmiclass}
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \bfseries {Classification} & \bfseries {BMI}\\
            \hline
            Underweight & \textless 20.0\\
            Normal weight/lean & 20.0$\sim$24.9\\
            Overweight & 25.0$\sim$29.9\\
            Obese & \textgreater 30.0\\
            \hline
        \end{tabular}
    \end{center}
\end{table}

\section{Download Data from publicly available source}
\label{sec:data}

@@

\subsection{The Cancer Genome Atlas (TCGA)}
\label{subsec:tcga}

@@

\subsection{International Cancer Genome Consortium (ICGC)}
\label{subsec:icgc}

RNA-seq data for the eight cancer types with sample BMI information available were downloaded from the ICGC database.
The RNA-seq data for each of the eight cancer types were imported into R.

\section{Validation of methodology used by \citet{Creighton2012}}
\label{sec:methodvalidation}

\subsection{Raw data from the \citet{Creighton2012} study}
\label{subsec:rawdatacr}

@@

(add data/sample composition -- race, grade, etc -- maybe in table form)

\subsection{Normalisation of raw \citet{Creighton2012} data}
\label{subsec:normcrdata}

@@

\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
\label{subsec:crobsgene}

799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (where? - i think the same place as microarray data).
The list of gene probes were imported into R for further analyses and validation of methodology.

\subsection{Differential gene expression analysis}
\label{subsec:deg}

All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample BMI status as defined by the WHO classification (see Table~\ref{tab:whobmiclass}).
The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in R.
The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.

The process above was repeated on the same data, with same group (obese and non-obese) after controlling for all other clinical variables that were not related to the sample BMI (age, sex, race, tumour grade, lymph node (LN) status, progesterone receptor (PR) status, and estrogen receptor (ER) status).
The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 between the two groups were identified.

To completely remove the effect of race on the analysis, the analysis was repeated using only the Caucasian samples.
This reduced the sample size from 103 samples to (70 something) samples (check sample size).
The data was controlled for all other clinical variables except the variables that were related to BMI (age, sex, tumour grade, LN status, PR status, and ER status).
Again, the top 799 differentially expressed genes between the two groups with p-value \textless{} 0.01 were identified.

\subsection{Comparison of \citet{Creighton2012} gene probes and identified gene probes}
\label{subsec:cfgeneprobes}

The obesity associated gene probes identified by \citet{Creighton2012} were compared with each of the three sets of gene probes identified from the differential gene expression analysis (see \ref{subsec:deg} \nameref{subsec:deg}).
The presence of any of the \citet{Creighton2012} genes were checked for in each of the gene sets identified at a p-value \textless{} 0.05, as well as the top 799 gene probes at p-Value \textless{} 0.01.

\subsection{Singular Value Decomposition}
\label{subsec:svd}

(need svd derivation and why I'm using it)

\subsection{Correlation between obesity associated metagene and sample BMI}
\label{subsec:metavsbmi}

In order to check whether the metagenes were actually specific to the BMI status of the samples, the metagenes were plotted against both the BMI statuses and the BMI values of the samples.
The metagenes were ranked for each sample, and a box plot and a scatter plot were plotted for sample BMI statuses and BMI values, respectively.

\section{Applying metagene to other cancer types}
\label{sec:metagene}

(How you could transfer the metagene to other cancer data)

\subsection{Transformation matrix}
\label{subsec:transmat}

(why transformation matrix is used)
(refer back to eqiations in the  svd section)

\section{Pathway Enrichment Analysis}
\label{sec:pathenrich}
